A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-03-19 graph_syntax.tex
added a new edge type

\section{MDAO Graph Primatives}
There are four node types:  
\begin{description}
\item[variable] represents scalar or array data
\item[model] responsible for mapping inputs to outputs
\item[expression] simple algebreic combinations of variables (such as for constraint of objective functions)
\item[driver] constrol structures capable of managing iteration
\end{description}

Only the variable, model, and expression nodes are allowed to be used in an MCG or FPF. 
All four nodes are required for any SPF. In addition to the three node types, there are two edge types:

\begin{description}
\item[fixed edge] A fixed edge may never be removed from the graph.
\item[free edge] A free edge may be removed removed from the graph.
\end{description}

A number of restrictions are placed on the model node with respect to the number and 
type of edges that can be connected to it: 
\begin{enumerate}
\item A model node can only have one edge directed to or from another model node.
\item A model node can only have fixed edges directed in or out.
\item A model node must have at least one edge directed in and at least one edge directed out.
\item If a variable node with an outgoing edge to a model node may not have any other outgoing edges.
\end{enumerate}

@@

Design variables, objectives, and constraints all share the property, that they are 
require external decisions to be made about a given system to identify them. Consider 
a notional problem given by the graph in Fig. [insert ref]a. There are three input variable 
nodes which, could potentially become design variables: w, x, and y. Assume that the 
output y is connected to the input y, yeilding Fig. [insert ref]b. Now only w and x could 
be design variables. We define a design variable to be any input variable node that does 
not have an incoming connection.

[[INSERT GRAPH HERE]]

>>>>

\section{MDAO Graph Primatives}
There are four node types:  
\begin{description}
\item[variable: ] represents scalar or array data
\item[model:] responsible for mapping inputs to outputs
\item[expression:] simple algebreic combinations of variables (such as for constraint of objective functions)
\item[driver:] constrol structures capable of managing iteration
\end{description}

In addition to the four node types, there are three edge types:

\begin{description}
\item[fixed edge:] Fundamental, non-changeable property of a problem, such as 
  connections between related variable nodes. These 
edges can never be removed from a graph. 
\item[explicit edge:] Explicit exchange of information between two nodes. No node 
  can have more than a single incomming explicit edge. These edges are free and
  can be added or removed from the graph. 
\item [implicit edge:] Implicit passing of information from a driver node to a 
  variable node. A single variable node can have many incomming implicit edges. These edges are 
  free and can be added or removed from the graph. 
\end{description}

The driver node and the implicit edge are only allowed to be pressent in SPF graphs. All 
other node and edge types can be present in in any of the three graphs, subject to a 
few restrictions: 
\begin{enumerate}
\item A model node can only have one edge directed to or from another model node.
\item A model node can only have fixed edges directed in or out.
\item A model node must have at least one edge directed in and at least one edge 
  directed out.
\item If a variable node has an outgoing edge to a model node then it may not have 
  any other outgoing edges.
\end{enumerate}

@@

Design variables, objectives, and constraints all require external decisions 
to be made about a given system to identify them. Consider 
a notional problem given by the graph in Fig. \ref{f:design vars} There are three input variables 
which, could potentially become design variables: $w$, $x$, and $y$. Because the output, $x$, of $A$
is connected to the input, $x$, of $B$ x can't actually be a design variable. The inputs $w$ and $y$
have no incomming edges and could each be selected as design variables.

\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
  \end{center}
  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
\end{figure}

More formally, all inputs that have no incomming fixed or explicit edges are defined
as design variables.
